---
title: "Class Project Draft"
author: "Anonymous"
date: "11/17/2024"
output: html_notebook
---

# Research Question
**"How have land temperatures changed over the past century across different countries, and what trends can we observe in global warming?"**

---

## Data Sources

**Primary Data Source:**  
- Dataset: *GlobalLandTemperaturesByCountry.csv*  
- Source: Kaggle, curated by Berkeley Earth.  
- Description: Historical land temperature records for countries worldwide. Each record contains the date, temperature, and country.  

**Secondary Data Source:**  
- Dataset: Meteorological data from the `climate` R package.  
- Description: Provides environmental variables like CO2 levels, precipitation, and temperature, adding complementary insights.

---
## Data Access and Inspection

### Importing Data
```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)

# Load primary dataset
temperature_data <- read_csv("/Users/vishallilani/Desktop/STAT 184/archive/GlobalLandTemperaturesByCountry.csv")

# Glimpse of the data
glimpse(temperature_data)

# Summary statistics
summary(temperature_data)
```

###Cleaning and Filtering
```{r}
recent_data <- temperature_data %>%
  filter(as.Date(dt) >= as.Date("1924-01-01"))

# Summary statistics for recent data
summary_stats <- recent_data %>%
  group_by(Country) %>%
  summarize(
    mean_temp = mean(AverageTemperature, na.rm = TRUE),
    max_temp = max(AverageTemperature, na.rm = TRUE),
    min_temp = min(AverageTemperature, na.rm = TRUE)
  )
summary_stats

```

###Exploring Global Trends
```{r}
sample_data <- temperature_data %>%
  sample_n(1000)

# Line graph showing temperature trends
ggplot(sample_data, aes(x = as.Date(dt), y = AverageTemperature)) +
  geom_line(alpha = 0.5, color = "blue") +
  labs(
    title = "Average Temperature Trends (Sampled Data)",
    x = "Date",
    y = "Average Temperature (째C)"
  ) +
  theme_minimal()

```

This graph shows fluctuating average temperatures over the years, with a notable increase in variability and values in recent decades, reflecting potential global warming trends.


###Comparing Regional Trends
```{r}
regional_trends <- recent_data %>%
  mutate(Region = case_when(
    Country %in% c("India", "China", "Japan") ~ "Asia",
    Country %in% c("United States", "Canada") ~ "North America",
    TRUE ~ "Other"
  )) %>%
  group_by(Region, dt) %>%
  summarize(mean_temp = mean(AverageTemperature, na.rm = TRUE))

# Visualization
ggplot(regional_trends, aes(x = as.Date(dt), y = mean_temp, color = Region)) +
  geom_line() +
  labs(
    title = "Temperature Trends by Region",
    x = "Date",
    y = "Mean Temperature (째C)",
    color = "Region"
  ) +
  theme_minimal()
```

This graph illustrates that different regions experience varied temperature trends. While Asia and North America show a consistent upward trend, the "Other" regions exhibit more variability.



###Global Temperature Distribution
```{r}

ggplot(recent_data, aes(x = Country, y = AverageTemperature)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Average Temperatures by Country",
    x = "Country",
    y = "Average Temperature (째C)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
The boxplot indicates a wide distribution of temperatures across countries, with some outliers showing extremely high or low average temperatures, emphasizing the diversity of climatic conditions globally.


###CO2 Levels vs. Temperature
```{r}

library(dplyr)
library(ggplot2)
library(lubridate)  


library(climate)
data("co2_demo", package = "climate")  


recent_data <- recent_data %>%
  mutate(Year = year(as.Date(dt)))  


co2_demo <- co2_demo %>%
  rename(Year = yy)

# Merge the datasets 
merged_data <- inner_join(recent_data, co2_demo, by = "Year")

# Create plot: CO2 levels vs. Land Temperature
ggplot(merged_data, aes(x = co2_avg, y = AverageTemperature)) +
  geom_point(alpha = 0.5, color = "blue") + 
  geom_smooth(method = "lm", color = "red") +  
  labs(
    title = "CO2 Levels vs. Land Temperature",
    x = "CO2 Levels (ppm)",
    y = "Average Land Temperature (째C)"
  ) +
  theme_minimal()
```

This scatter plot with a regression line highlights a positive relationship between CO2 levels and average land temperatures, supporting the theory that increased CO2 levels are associated with global warming.


