---
title: "Class Project Draft"
author: "Anonymous"
date: "11/17/2024"
output: html_notebook
---

# Research Question
**"How have land temperatures changed over the past century across different countries, and what trends can we observe in global warming?"**

---

## Data Sources

**Primary Data Source:**  
- Dataset: *GlobalLandTemperaturesByCountry.csv*  
- Source: Kaggle, curated by Berkeley Earth.  
- Description: Historical land temperature records for countries worldwide. Each record contains the date, temperature, and country.  

**Secondary Data Source:**  
- Dataset: Meteorological data from the `climate` R package.  
- Description: Provides environmental variables like CO2 levels, precipitation, and temperature, adding complementary insights.

---
## Data Access and Inspection

### Importing Data
```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(readr)

# Load primary dataset
temperature_data <- read_csv("/Users/vishallilani/Desktop/STAT 184/archive/GlobalLandTemperaturesByCountry.csv")

# Glimpse of the data
glimpse(temperature_data)

# Summary statistics
summary(temperature_data)
```

###Cleaning and Filtering
```{r}
recent_data <- temperature_data %>%
  filter(as.Date(dt) >= as.Date("1924-01-01"))

# Summary statistics for recent data
summary_stats <- recent_data %>%
  group_by(Country) %>%
  summarize(
    mean_temp = mean(AverageTemperature, na.rm = TRUE),
    max_temp = max(AverageTemperature, na.rm = TRUE),
    min_temp = min(AverageTemperature, na.rm = TRUE)
  )
summary_stats

```

###Exploring Global Trends
```{r}
sample_data <- temperature_data %>%
  sample_n(1000)

# Line graph showing temperature trends
ggplot(sample_data, aes(x = as.Date(dt), y = AverageTemperature)) +
  geom_line(alpha = 0.5, color = "blue") +
  labs(
    title = "Average Temperature Trends (Sampled Data)",
    x = "Date",
    y = "Average Temperature (°C)"
  ) +
  theme_minimal()

```

This graph shows fluctuating average temperatures over the years, with a notable increase in variability and values in recent decades, reflecting potential global warming trends.



